---
title: "SpotifyEDA"
output: html_notebook
---

#import libraries
```{r}
library(tidyverse)
library(ggplot2)
```

#import csvs
```{r}
fullmaster <- read_csv('../data/fullmasterfinal.csv')
covid <- read_csv('../data/covidmaster.csv')
```
# avg danceability distribution each month
```{r}
fullmaster |>
  group_by(year, month) |>
  summarize(avgdanceability = mean(danceability, na.rm = TRUE))
```

# avg energy distribution each month
```{r}
fullmaster |>
  group_by(year, month) |>
  summarize(avgenergy = mean(energy, na.rm = TRUE))
```



# avg loudness distribution each month
```{r}
fullmaster |>
  group_by(year, month) |>
  summarize(avgloudness = mean(loudness, na.rm = TRUE))
```


# plot avg valence each month grouped by year
```{r}
avgdf <- fullmaster |>
  group_by(year, month) |>
  summarize(Average = mean(valence, na.rm = TRUE)) |>
  mutate(month = factor(month), year = factor(year))

avgplot <- avgdf|> 
  ggplot(aes(month, Average, group = year, color = year)) +
  geom_line()
print(avgplot)
```
#anova for significance. none, aww
```{r}
avg_anova <- aov(avgvalence ~ year + Error(month/year), data = avgdf)
anova_output <- summary(avg_anova)
anova_output[[2]][[1]][1,5]
```
#linear regression for significance between years, to use all data instead of just the average
```{r}
linear_model <- lm(liveness ~ factor(year)+month, data= fullmaster)
summary(linear_model)
```
#checking  if at least one month is significantly different
```{r}
reduced_model <- lm(liveness ~ month, data= fullmaster)
summary(reduced_model)
```

```{r}
full_model <- lm(liveness ~ month*factor(year), data= fullmaster)
summary(full_model)
```
#anova on lm
```{r}
lm_anova <- anova(reduced_model, full_model)
lm_anova
```
#t-test on the anova adjusted for months(p-value*12 <0.05?)
```{r}

months <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
p_values <- c()
selected_month <- "07"
selected_variable <- "liveness"

values2019 <- fullmaster |>
filter(month == selected_month, year == "2019") |>
pull(selected_variable)

values2020 <- fullmaster |>
filter(month == selected_month, year == "2020") |>
pull(selected_variable)

(t.test(values2019, values2020)$p.value)*12
```


```{r}
months <- c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
p_values <- c()

for (selected_month in months){
  selected_variable = "energy"
  values2019 <- fullmaster |>
    filter(month == selected_month, year == "2019") |>
    pull(selected_variable)
  values2020 <- fullmaster |>
    filter(month == selected_month, year == "2020") |>
    pull(selected_variable)
    p_values = c(p_values, (t.test(values2019, values2020)$p.value)*12)}

tibble(month = months, pvalue = p_values)
```








#group COVID data by month
```{r}
monthlycovid <- covid |>
  group_by(month) |>
  summarize(deaths = sum(deaths), allcausedeaths = sum(allcausedeaths), percentexpected = mean(percentexpected))
```

#plot COVID deaths
```{r}
monthlycovid |>
  mutate(month = as.numeric(month)) |>
  ggplot(aes(month, deaths)) +
  geom_line() +
  scale_x_continuous(breaks = 1:12)
```

```{r}
fullmaster |>
  group_by(year, month, key) |>
  mutate(key = c("C", "C#", "D", "D#", "E", "F", "F#", "G", "G#","A", "A#", "B")[key+1], year = factor(year)) |>
  mutate(key = factor(key, levels = c("C", "C#", "D", "D#", "E", "F", "F#", "G", "G#","A", "A#", "B"))) |>
  drop_na(key) |>
  summarise(count = n()) |>
  ggplot(aes(x=year, y= count, fill= key))+
  geom_bar(stat = "identity", position = "fill", color= "black", size= 0.1, width= 0.5)+
  facet_wrap(~month)+
  labs(x="Month", y="Number of Songs") 
  #guide_legend = 
  #guides(fill=guide_legend(title="Key"))
```
#check for significance(all are significant??)
```{r}
key_contingencies <- fullmaster |>
  drop_na(key) |>
  count(year, month, key) |>
  spread(year, n, fill = 0)
key_chi_square <- key_contingencies |>
  group_split(month) |>
  map(~ chisq.test(select(.x, -month))) |>
  set_names(unique(fullmaster$month))
key_chi_square[[1]][3]
```
#determine dominant key each month
```{r}
dominant_keys <- fullmaster |>
  group_by(year, month, key, mode) |>
  summarise(count = n()) |>
  slice(which.max(count))
dom_key_clean <- dominant_keys |>
  group_by(year, month) |>
  slice(which.max(count))
dom_key_clean

```
```{r}

```


#song key examples
```{r}
fullmaster |>
  filter(key == 10, mode ==0)
```






Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

